---
title: "Untitled"
author: "Jens Hüsers"
date: "13 11 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Basis Adress um alle Devices auf dem Server anzufragen:
http://fhir.hl7fundamentals.org/baseDstu3/Device

When we use a resource of type `Reference` in FHIR there are two main ways how a reference can be made.

The first one is the `Reference.reference` data field.
This field contains

In order to query a device of a specific manufacturer we can use the `owner` reference in the resource device.

The owner can have different reference types.

1. the id of the manufacturer on the FHIR server
2. the business identifier given by `owner.identifier`

We can query the reference 

`http://fhir.hl7fundamentals.org/baseDstu3/Device?organization=29347`

Right now I have some problems quering a device resource when the organization reference is a business identifier.


# Create and update update a FHIR resource

In this post I want to demonstrate how to create and update a FHIR resource.

I am going to use the `PUT` operation for this task.
`PUT` will create a new resource on the server if this resource is not yet on the server.
In opposite to the `POST` operation which expects the `[base-url]` with the resource type to create, e.g. `Patient`, `PUT` expects and additional identifier.
Therefore, the final `PUT` operation is composted by the `[base-url]` part, the `resource-type` and an `identifier` value.

When we first `PUT` a resource on the server, we don't know the local resource `id` because it has not been created yet.
So I will use the business identifier of the patient by using the element [`identifier`](https://www.hl7.org/fhir/patient-definitions.html#Patient.identifier) of the patient resource.



```{r eval=FALSE}

<Patient>
  ... 
  <identifier>
    <system value="mypatientregistry.com"/>
    <value value="23qcqdstq3q2e"/>
  </identifier>
  ...
</Patient>
  
```

So I can compose the url to create/ update my resource:

`[base-ur]/Patient?identifier=mypatientregistry.com|23qcqdstq3q2e`

So if my resource is not on the server the resource will created and the server returns the `201 OK` code indicating that the resource was created.
If the resource exists on the server and I made changes to the resource, e.g. editing the patient name, the resource is updated and the server returns the code `200 OK`.

# Examples

### Step 1 - Create Resource

I constructed a patient example resource with an business identifier.

```{r eval=FALSE}

<Patient>
  <identifier>
    <system value="www.mypatientidentifier.com/ids"/>
    <value value="23894734287402eu"/>
  </identifier>
  <name>
    <family value="Jens"/>
    <given value="Spahn"/>
  </name>
  <address>
    <line value="Hauptstraße 16"/>
    <city value="Berlin"/>
    <state value="Berlin"/>
    <country value="France"/>
  </address>
</Patient>
  
```

I create the patient resource using the following url:

`http://hapi.fhir.org/baseDstu3/Patient?identifier=www.mypatientidentifier.com/ids|8qww76emweqs1`

The server returns the code `201 Creates` which indicates that the resource was created.

### Step 2 - Update Resource

I edited the previous resource because I accidentially inserted the wrong country in the `address` element.

In this case I use the following resource to update the resource.

```{r eval=FALSE}

<Patient>
  <identifier>
    <system value="www.mypatientidentifier.com/ids"/>
    <value value="23894734287402eu"/>
  </identifier>
  <name>
    <family value="Jens"/>
    <given value="Spahn"/>
  </name>
  <address>
    <line value="Hauptstraße 16"/>
    <city value="Berlin"/>
    <state value="Berlin"/>
    <country value="Germany"/>
  </address>
</Patient>
  
```

The server returns `200 OK` code and we can see that the `versionId` value incrementally increased and has the value 2.

It is important to send the complete resource and not only the element you want to update as in the example below.

```{r eval=FALSE}

<Patient>
  <identifier>
    <system value="www.mypatientidentifier.com/ids"/>
    <value value="23894734287402eu"/>
  </identifier>
  <address>
    <country value="Germany"/>
  </address>
</Patient>
  
```

In this case the server again returns `200 OK` but this time all other elements of the resource are dropped and the only element left in the updated resource is the `address.country` element.



1. Id in die Resource geben
2. Diese Id wird überschrieben
3. Die neue Id aus der Outcome Resource enthält die überschriebene Id